Simple repression draft

### 2020-05-30 doc/section_04_bayesian_inference.tex
MRM pass through section 04. No big changes

In this section of the paper, we continue our program of providing
one complete description of the entire broad sweep of studies
that have been made in the context of the repressor-operator model,
dating all the way back to the original work of Jacob and Monod
and including the visionary quantitative work of M\"{u}ller-Hill and collaborators
and up to more recent studies.   In addition, the aim is to reconcile the 
equilibrium and non-equilibrium perspectives that have been brought
to bear on this problem. From Section~\ref{section_02_means}, this reconciliation
depends on a key quantitative question as codified by Eq.~\ref{eq:deltaFR_eq_noneq_equiv}: does the free energy of repressor binding, as described in the equilibrium models and indirectly inferred from gene expression measurements, agree with the corresponding values of repressor binding and unbinding rates in the non-equilibrium picture, measured or inferred more directly? In this section we tackle the statistical inference problem of inferring these repressor rates from mRNA FISH data. But before we can turn to the full case of simple repression, we must choose an appropriate model of the constitutive promoter and infer the parameter values in that model. This is the problem we address first.

@@

As detailed in Appendix~\ref{sec:bayesian} this particular choice of prior is known
as the \textit{conjugate} prior for a Poisson likelihood. This type of priors
have nice properties to work with, in particular a conjugate prior allows us to
get a closed form for the posterior distribution $p(\lambda \mid D)$. In
particular it can be shown that for a Poisson distribution likelihood with its
Gamma distribution conjugate prior, the posterior distribution is also a Gamma
distribution ~\cite{Gelman2013}. In particular the two parameters $\alpha'$ and
$\beta'$ for this posterior distribution take the form $\alpha' = \alpha +
\bar{m} N$ and $\beta' = \beta + N$, where we defined the sample mean $\bar{m} =
\frac{1}{N}\sum_k m_k$ for notational convenience, and $N$ is the number of
cells in our dataset. Furthermore, given that $N$ is $\mathcal{O}(10^3)$ and
$\langle m\rangle \gtrsim 0.1$ for all promoters measured in~\cite{Jones2014}
our data easily overwhelms the choice of prior, and allows us to approximate the
Gamma distribution with a Gaussian distribution with mean $\bar{m}$ and variance
$\bar{m} / N$ with marginal errors. As an example with real numbers, for the
\textit{lacUV5} promoter, Jones et.\ al~\cite{Jones2014} measured 2648 cells
with an average mRNA count per cell of $\bar{m} \approx 18.7$. For this case our
posterior distribution $P(\lambda \mid D)$ would be a Gaussian distribution with
mean $\mu = 18.7$, and a standard deviation $\sigma \approx 0.08$. This suggests
we have inferred our model's one parameter to a precision of order 1\%.

@@

Given this result, how are the two coarse-grainings related? A quick estimate
can shed some light. Consider for instance the \textit{lacUV5} promoter, which
we see from Figure~\ref{fig:constit_post_full}(A) has $k_i/\gamma \sim b \sim
\text{few}$, from Figure~\ref{fig:constit_post_full}(B) has $\langle m \rangle
\sim 20$, and from~\cite{Brewster2012} has $\beta\Delta\epsilon_P \sim - 6.5$.
Further we generally assume $P/N_{NS} \sim 10^{-3}$ since
$N_{NS}\approx4.6\times10^6$ and $P\sim10^3$. After some guess-and-check with
these values, one finds the only association that makes dimensional sense and
produces the correct order-of-magnitude for the known parameters is to take
\begin{equation}
\frac{k_i}{\gamma} = \frac{P}{N_{NS}} \exp(-\beta\Delta\varepsilon_P)
\label{eq:bursty_equil_corresp1}
\end{equation}
and
\begin{equation}
b = \frac{r}{\gamma}.
\label{eq:bursty_equil_corresp2}
\end{equation}
Figure~\ref{fig:constit_post_full}(D) shows that this linear scaling between
$\ln k_i$ and $-\beta\Delta\varepsilon_P$ is approximately true for all 18
constitutive promoters considered. The plotted line is simply
Eq.~\ref{eq:bursty_equil_corresp1} and assumes $P\approx 5000$.

@@

Formally now, denote the set of seven repressor rates to be inferred as
\begin{equation}
\vect{k} =\{k_{Oid}^-, k_{O1}^-, k_{O2}^-,
k_{0.5}^+, k_{1}^+, k_{2}^+, k_{10}^+\},
\end{equation}
where subscripts for dissociation rates $k^-$  indicate the different repressor
binding sites, and subscripts for association rates $k^+$ indicate the aTc
concentration at which the repressor itself was induced. We emphasize that for
this particular dataset the repressor copy numbers are not known directly, so we
label their association rates by the concentration of a small molecule inducer
aTc which indirectly controls the transcription of $\textit{lacI}$, as
described in detail in~\cite{Jones2014}. Also note
that the authors of~\cite{Jones2014} report estimates of LacI copy number per
cell rather than direct measurements. However, these estimates were made
\textit{assuming} the validity of the equilibrium models in
Figure~\ref{fig1:means_cartoons}, and since \textit{testing} these models is our
present goal, it would be circular logic if we were to make the same assumption.
Therefore we will make no assumptions about the LacI copy number for given aTc
concentrations.

Having stated the problem, Bayes theorem reads simply
\begin{equation}
p(\vect{k}, k_i, b \mid D)
\propto
p(D \mid\vect{k}, k_i, b) p(\vect{k}, k_i, b),
\end{equation}
where $D$ is again the set of all $N$ observed single-cell mRNA counts
across the various conditions. We assume that individual single-cell
measurements are independent so that the likelihood factorizes as
\begin{equation}
p(D \mid\vect{k}, k_i, b)
= \prod_{j=1}^N p(m\mid \vect{k}, k_i, b)
= \prod_{j=1}^N p(m\mid k_j^+, k_j^-, k_i, b)
\end{equation}
where $k_j^\pm$ represent the appropriate binding and unbinding
rates out of $\vect{k}$ for the $j$-th measured cell. The probability
$p(m\mid k_j^+, k_j^-, k_i, b)$ appearing in the last expression
is exactly Eq.~\ref{eq:p_m_bursty+rep_appdx}, the steady-state
distribution for our bursty model with repression derived in
Section~\ref{sec:gen_fcn_appdx}, which for completeness we reproduce here as
\begin{equation}
\begin{split}
p(m \mid k_R^+, k_R^-, k_i, b) = & ~\frac{
        \Gamma(\alpha + m)\Gamma(\beta + m)\Gamma(k_R^+ + k_R^-)
        }
        {
        \Gamma(\alpha)\Gamma(\beta)\Gamma(k_R^+ + k_R^- + m)
        }
\frac{b^m}{m!}
\\
&\times {_2F_1}(\alpha+m, \beta+m, k_R^++k_R^-+m; -b).
\end{split}
\label{eq:p_m_bursty+rep}
\end{equation}
where $\alpha$ and $\beta$, defined for notational convenience, are
\begin{align}
\begin{split}
\alpha &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ + \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right)
\\
\beta &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ - \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right).
\end{split}
\end{align}

@@

\mmnote{These next 2 paragraphs should be SI, I think. Leaving here for
now to avoid a merge conflict w/ MRM.}

We found that fitting a single operator/aTc concentration at a
time with a single binding and unbinding rate did not yield a
stable inference for most of the possible operator/aTc
combinations. In other words, a single dataset could not
independently resolve the binding and unbinding rates, only their
ratio as set by the mean fold-change in
Figure~\ref{fig1:means_cartoons}. Only by making the assumption
of a single unique binding rate for each repressor copy number
and a single unique unbinding rate for each binding site, as done
in Figure~\ref{fig4:repressed_post_full}A, was it possible to
independently resolve the rates and not merely their ratios.

We also note that we found it necessary to exclude the very
weakly and very strongly repressed datasets from Jones et.\
al.~\cite{Jones2014}. In both cases there was, in a sense, not
enough information in the distriubtions for our inference
algorithm to extract, and their inclusion simply caused problems
for the MCMC sampler without yielding any new insight. For the
strongly repressed data (Oid, 10~ng/mL aTc), with $>$ 95\% of cells
with zero mRNA, there was quite literally very little data from
which to infer rates. And the weakly repressed data, all with the
repressor binding site O3, had an unbinding rate so fast that the
sampler essentially sampled from the prior; the likelihood had
negligible influence, meaning the data was not informing the
sampler in any meaninful way, so no inference was possible.

@@

Figure~\ref{fig4:repressed_post_full}(B) shows both comparisons,
with the top panel comparing to equilbrium binding energies
and the bottom panel comparing to single-molecule measurements.
First consider the top panel and its comparison between repressor
kinetic rates and binding energies.
As described in section~\ref{section_02_means}, if the
equilibrium binding energies from~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018} indeed are the physically real binding
energies we believe them to be, then they should be related to
the repressor kinetic rates via
Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, which we restate here,
\begin{equation}
\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-).
\label{eq:deltaFR_eq_noneq_equiv_repeat}
\end{equation}
Assuming mass action
kinetics implies that $k_R^+$ is proportional to repressor copy
number $R$, or more precisely, it can be thought of as repressor
copy number times some intrinsic per molecule association rate.
But since $R$ is not directly known for our data
from~\cite{Jones2014}, we cannot use this equation directly.
Instead we can consider two different repressor binding sites and
compute the \textit{difference} in binding energy between them,
since this difference depends only on the unbinding rates and not
on the binding rates. This can be seen by evaluating
Eq.~\ref{eq:deltaFR_eq_noneq_equiv_repeat} for two different
repressor binding sites, labeled (1) and (2), but with the same
repressor copy number $R$, and taking the difference to find
\begin{equation}
\Delta F_R^{(1)} - \Delta F_R^{(2)}
= \beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= - \log(k_R^+/k_1^-) + \log(k_R^+/k_2^-),
\end{equation}
or simply
\begin{equation}
\beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= \log(k_2^-/k_1^-).
\end{equation}
The left and right hand sides of this equation are exactly the horizontal
and vertical axes of the top panel of Figure~\ref{fig4:repressed_post_full}.
Since we inferred rates for three repressor binding sites (O1, O2, and Oid),
there are only two independent differences that can be constructed, and we
arbitrarily chose to plot O2-O1 and O1-Oid in
Figure~\ref{fig4:repressed_post_full}(B).
Numerically, we compute values of $k_{O1}^- / k_{Oid}^-$ and
$k_{O2}^- / k_{O1}^-$ directly from our full posterior samples,
which conveniently provides uncertainties as well, as detailed in
Appendix~\ref{sec:bayesian}.
\mmnote{Explain in appendix how marginalization with sampling makes this
``trivial'' to compute from samples of the joint posterior, b/c
sampling makes marginalization trivial.}
We then compare these ratios of rates to the binding energy differences
$\Delta\epsilon_{O1} - \Delta\epsilon_{Oid}$ and from
$\Delta\epsilon_{O2} - \Delta\epsilon_{O1}$ as computed from the values
from both~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}.
Three of the four values are within $\sim0.5 k_BT$ of the diagonal
representing perfect agreement, which is comparable to the
$\sim\text{few}\times0.1 k_BT$ variability between the independent
determinations of the same quantities between~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}. The only outlier involves Oid measurements
from~\cite{Razo-Mejia2018}, and as the authors of~\cite{Razo-Mejia2018} note,
this is a difficult measurement of low fluorescence signal against
high background since Oid represses so strongly. We are therefore
inclined to regard the failure of this point to fall near the
diagonal as a testament to the difficulty of the measurement and
not as a failure of our theory.
\mmnote{Does this seem a fair characterization?   - RP: thanks, that was very helpful
for me.}

@@

Since we do not have access to repressor copy number for either
the FISH data from~\cite{Jones2014} or the single-molecule data
from~\cite{Hammar2014}, we cannot make an apples-to-apples
comparison of association rates $k_R^+$. Further, while Hammar
et.\ al.\ directly measure the dissociation rates $k_R^-$, our
inference procedure returns $k_R^-/\gamma$, i.e., the repressor
dissociation rate nondimensionalized by the mRNA degradation rate
$\gamma$. So to make the comparison, we must make an assumption
for the value of $\gamma$ since it was not directly measured. For
most mRNAs in \textit{E.\ coli}, quoted values for the typical mRNA
lifetime $\gamma^{-1}$ range between about 2.5~min~\cite{Chen2015}
to 8~min. We chose $\gamma^{-1} = 3$~min and $\gamma^{-1} =
5$~min as representative values and plot a comparison of
$k_{O1}^-$ and $k_{Oid}^-$ from our inference with corresponding values
reported in~\cite{Hammar2014} for both these choices of $\gamma$.

The degree of quantitative agreement in the lower panel of
Figure~\ref{fig4:repressed_post_full}B clearly depends on the
precise choice of $\gamma$. Nevertheless we find this comparison very satisfying, when two wildly different approaches to a measurement of the same quantity yield broadly compatible results. We emphasize the agreement between our rates
and the rates reported in~\cite{Hammar2014} for any reasonable
$\gamma$: values differ by at most a factor of 2 and possibly
agree to within our uncertainties of 10-20\%. From this we feel
confident asserting that the parameters we have inferred from
Jones et.\ al.'s FISH data do in fact correspond to repressor
binding and unbinding rates, and therefore our conclusions on the
agreement of these rates with binding energies
from~\cite{Garcia2011a} and~\cite{Razo-Mejia2018} are valid.

@@

The details are deferred to Appendix~\ref{sec:bayesian}, and here
we only attempt to summarize the intuitive ideas, as detailed at
greater length by Jaynes~\cite{Jaynes2003} as well as Gelman and
coauthors~\cite{Gelman2013,Gelman2013a}. From our samples of the
posterior distribution, plotted in
Figure~\ref{fig4:repressed_post_full}A, we generate many
replicate data using a random number generator.
In Figure~\ref{fig4:repressed_post_full}C, we plot ECDFs of the
middle 95\% quantiles of these replicate data with the actual
experimental data from Jones et.\ al.~\cite{Jones2014} overlaid,
covering all ten experimental conditions spanning repressor
binding sites and copy numbers (as well as the constitutive
baseline UV5).

The purpose of Figure~\ref{fig4:repressed_post_full}C is simply a
graphical, qualitative assessment of the model: do the
experimental data systematically disagree with the simulated
data, which would suggest that our model is missing important
features? A further question is not just whether there is a
detectable difference between simulated and experimental data,
but whether this difference is likely to materially affect the
conclusions we draw from the posterior in
Figure~\ref{fig4:repressed_post_full}A. More rigorous and
quantitative statistical tests are possible~\cite{Gelman2013},
but their quantitativeness does not necessarily make them more
useful. As stated in~\cite{Gelman2013a}, we often find this
graphical comparison more enlightening because it better engages
our intuitition for the model, not merely telling \textit{if} the
model is wrong but suggesting \textit{how} the model may be
incomplete.

Our broad brush takeaway from
Figure~\ref{fig4:repressed_post_full}C is overall of good
agreement. There some oddities, in particular the long tails in
the data for Oid, 1~ng/mL, and O2, 0.5~ng/mL. The latter is
especially odd since it extends beyond the tail of the
unregulated UV5 distribution. This is a relatively small number
of cells, however, so whether this is a peculiarity of the
experimental data, a statistical fluke of small numbers, or a
real biological effect is unclear. It is conceivable that there
is some very slow timescale switching dynamics that could cause
this bimodality, although it is unclear why it would only appear
for specific repressor copy numbers. There is also a small offset
between experiment and simulation for O2 at the higher repressor
copy numbers, especially at 2~ng/mL. We suspect this may be
caused by the very large number of zero mRNA counts for Oid,
2~ng/mL skewing it's partner datasets through the shared
association rate, and so we are not too concerned by this
discrepancy either. None of these fairly minor quibbles cause us
to seriously doubt the overall correctness of our model, which
further validates its use to compare the equilbrium models'
binding energies to the nonequilibrium models' repressor
kinetics, as we originally set out to do.

>>>>

In this section of the paper, we continue our program of providing one complete
description of the entire broad sweep of studies that have been made in the
context of the repressor-operator model, dating all the way back to the original
work of Jacob and Monod and including the visionary quantitative work of
M\"{u}ller-Hill and collaborators~\cite{Oehler1990} and up to more recent
studies \cite{Garcia2011a}. In addition, the aim is to reconcile the equilibrium
and non-equilibrium perspectives that have been brought to bear on this problem.
From Section~\ref{section_02_means}, this reconciliation depends on a key
quantitative question as codified by Eq.~\ref{eq:deltaFR_eq_noneq_equiv}: does
the free energy of repressor binding, as described in the equilibrium models and
indirectly inferred from gene expression measurements, agree with the
corresponding values of repressor binding and unbinding rates in the
non-equilibrium picture, measured or inferred more directly? In this section we
tackle the statistical inference problem of inferring these repressor rates from
mRNA FISH data. But before we can turn to the full case of simple repression, we
must choose an appropriate model of the constitutive promoter and infer the
parameter values in that model. This is the problem we address first.

@@

As detailed in Appendix~\ref{sec:bayesian} this particular choice of prior is
known as the \textit{conjugate} prior for a Poisson likelihood. This type of
priors have nice properties to work with, in particular a conjugate prior allows
us to get a closed form for the posterior distribution $p(\lambda \mid D)$. For
our case it can be shown that for a Poisson distribution likelihood with its
Gamma distribution conjugate prior, the posterior distribution is also a Gamma
distribution ~\cite{Gelman2013}. Specifically the two parameters $\alpha'$ and
$\beta'$ for this posterior distribution take the form $\alpha' = \alpha +
\bar{m} N$ and $\beta' = \beta + N$, where we defined the sample mean $\bar{m} =
\frac{1}{N}\sum_k m_k$ for notational convenience, and $N$ is the number of
cells in our dataset. Furthermore, given that $N$ is $\mathcal{O}(10^3)$ and
$\langle m\rangle \gtrsim 0.1$ for all promoters measured in~\cite{Jones2014}
our data easily overwhelms the choice of prior, and allows us to approximate the
Gamma distribution with a Gaussian distribution with mean $\bar{m}$ and variance
$\bar{m} / N$ with marginal errors. As an example with real numbers, for the
\textit{lacUV5} promoter, Jones et.\ al~\cite{Jones2014} measured 2648 cells
with an average mRNA count per cell of $\bar{m} \approx 18.7$. For this case our
posterior distribution $P(\lambda \mid D)$ would be a Gaussian distribution with
mean $\mu = 18.7$, and a standard deviation $\sigma \approx 0.08$. This suggests
we have inferred our model's one parameter to a precision of order 1\%.

@@

Given this result, how are the two coarse-grainings related? A quick estimate
can shed some light. Consider for instance the \textit{lacUV5} promoter, which
we see from Figure~\ref{fig:constit_post_full}(A) has $k_i/\gamma \sim b \sim
\text{few}$, from Figure~\ref{fig:constit_post_full}(B) has $\langle m \rangle
\sim 20$, and from~\cite{Brewster2012} has $\beta\Delta\varepsilon_P \sim -
6.5$. Further we generally assume $P/N_{NS} \sim 10^{-3}$ since
$N_{NS}\approx4.6\times10^6$ and $P\sim10^3$. After some guess-and-check with
these values, one finds the only association that makes dimensional sense and
produces the correct order-of-magnitude for the known parameters is to take
\begin{equation}
\frac{k_i}{\gamma} = \frac{P}{N_{NS}} \exp(-\beta\Delta\varepsilon_P)
\label{eq:bursty_equil_corresp1}
\end{equation}
and
\begin{equation}
b = \frac{r}{\gamma}.
\label{eq:bursty_equil_corresp2}
\end{equation}
Figure~\ref{fig:constit_post_full}(D) shows that this linear scaling between
$\ln k_i$ and $-\beta\Delta\varepsilon_P$ is approximately true for all 18
constitutive promoters considered. The plotted line is simply
Eq.~\ref{eq:bursty_equil_corresp1} and assumes $P\approx 5000$.

@@

Formally now, denote the set of seven repressor rates to be inferred as
\begin{equation}
\vect{k} =\{k_{Oid}^-, k_{O1}^-, k_{O2}^-,
k_{0.5}^+, k_{1}^+, k_{2}^+, k_{10}^+\},
\end{equation}
where subscripts for dissociation rates $k^-$  indicate the different repressor
binding sites, and subscripts for association rates $k^+$ indicate the aTc
concentration at which the repressor itself was induced. We emphasize that for
this particular dataset the repressor copy numbers are not known directly, so we
label their association rates by the concentration of a small inducer molecule 
aTc which indirectly controls the transcription of $\textit{lacI}$, as
described in detail in~\cite{Jones2014}. Also note
that the authors of~\cite{Jones2014} report estimates of LacI copy number per
cell rather than direct measurements. However, these estimates were made
\textit{assuming} the validity of the equilibrium models in
Figure~\ref{fig1:means_cartoons}, and since \textit{testing} these models is our
present goal, it would be circular logic if we were to make the same assumption.
Therefore we will make no assumptions about the LacI copy number for given aTc
concentrations.

Having stated the problem, Bayes theorem reads simply
\begin{equation}
p(\vect{k}, k_i, b \mid D)
\propto
p(D \mid\vect{k}, k_i, b) p(\vect{k}, k_i, b),
\end{equation}
where $D$ is again the set of all $N$ observed single-cell mRNA counts
across the various conditions. We assume that individual single-cell
measurements are independent so that the likelihood factorizes as
\begin{equation}
p(D \mid\vect{k}, k_i, b)
= \prod_{j=1}^N p(m\mid \vect{k}, k_i, b)
= \prod_{j=1}^N p(m\mid k_j^+, k_j^-, k_i, b)
\end{equation}
where $k_j^\pm$ represent the appropriate binding and unbinding
rates out of $\vect{k}$ for the $j$-th measured cell. The probability
$p(m\mid k_j^+, k_j^-, k_i, b)$ appearing in the last expression
is exactly Eq.~\ref{eq:p_m_bursty+rep_appdx}, the steady-state
distribution for our bursty model with repression derived in
Section~\ref{sec:gen_fcn_appdx}, which for completeness we reproduce here as
\begin{equation}
\begin{split}
p(m \mid k_R^+, k_R^-, k_i, b) = & ~\frac{
        \Gamma(\alpha + m)\Gamma(\beta + m)\Gamma(k_R^+ + k_R^-)
        }
        {
        \Gamma(\alpha)\Gamma(\beta)\Gamma(k_R^+ + k_R^- + m)
        }
\frac{b^m}{m!}
\\
&\times {_2F_1}(\alpha+m, \beta+m, k_R^++k_R^-+m; -b).
\end{split}
\label{eq:p_m_bursty+rep}
\end{equation}
where $_2F_1$ is the confluent hypergeometric function of the second kind,
$\alpha$ and $\beta$, defined for notational convenience, are
\begin{align}
\begin{split}
\alpha &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ + \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right)
\\
\beta &= \frac{1}{2}
\left(k_i+k_R^-+k_R^+ - \sqrt{(k_i+k_R^-+k_R^+)^2 - 4k_i k_R^-}\right).
\end{split}
\end{align}

@@

@@

Figure~\ref{fig4:repressed_post_full}(B) shows both comparisons, with the top
panel comparing to equilibrium binding energies and the bottom panel comparing
to single-molecule measurements. First consider the top panel and its comparison
between repressor kinetic rates and binding energies. As described in
section~\ref{section_02_means}, if the equilibrium binding energies
from~\cite{Garcia2011a} and~\cite{Razo-Mejia2018} indeed are the physically real
binding energies we believe them to be, then they should be related to the
repressor kinetic rates via Eq.~\ref{eq:deltaFR_eq_noneq_equiv}, which we
restate here,
\begin{equation}
\Delta F_R = \beta\Delta\varepsilon_R - \log(R/N_{NS})
        = - \log(k_R^+/k_R^-).
\label{eq:deltaFR_eq_noneq_equiv_repeat}
\end{equation}
Assuming mass action kinetics implies that $k_R^+$ is proportional to repressor
copy number $R$, or more precisely, it can be thought of as repressor copy
number times some intrinsic per molecule association rate. But since $R$ is not
directly known for our data from~\cite{Jones2014}, we cannot use this equation
directly. Instead we can consider two different repressor binding sites and
compute the \textit{difference} in binding energy between them, since this
difference depends only on the unbinding rates and not on the binding rates.
This can be seen by evaluating Eq.~\ref{eq:deltaFR_eq_noneq_equiv_repeat} for
two different repressor binding sites, labeled (1) and (2), but with the same
repressor copy number $R$, and taking the difference to find
\begin{equation}
\Delta F_R^{(1)} - \Delta F_R^{(2)}
= \beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= - \log(k_R^+/k_1^-) + \log(k_R^+/k_2^-),
\end{equation}
or simply
\begin{equation}
\beta\Delta\varepsilon_1 - \beta\Delta\varepsilon_2
= \log(k_2^-/k_1^-).
\end{equation}
The left and right hand sides of this equation are exactly the horizontal and
vertical axes of the top panel of Figure~\ref{fig4:repressed_post_full}. Since
we inferred rates for three repressor binding sites (O1, O2, and Oid), there are
only two independent differences that can be constructed, and we arbitrarily
chose to plot O2-O1 and O1-Oid in Figure~\ref{fig4:repressed_post_full}(B).
Numerically, we compute values of $k_{O1}^- / k_{Oid}^-$ and $k_{O2}^- /
k_{O1}^-$ directly from our full posterior samples, which conveniently provides
uncertainties as well, as detailed in Appendix~\ref{sec:bayesian}.
\mmnote{Explain in appendix how marginalization with sampling makes this
``trivial'' to compute from samples of the joint posterior, b/c sampling makes
marginalization trivial.} We then compare these log ratios of rates to the
binding energy differences $\Delta\varepsilon_{O1} - \Delta\varepsilon_{Oid}$
and from $\Delta\varepsilon_{O2} - \Delta\varepsilon_{O1}$ as computed from the
values from both~\cite{Garcia2011a} and~\cite{Razo-Mejia2018}. Three of the four
values are within $\sim0.5 k_BT$ of the diagonal representing perfect agreement,
which is comparable to the $\sim\text{few}\times0.1 k_BT$ variability between
the independent determinations of the same quantities between~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018}. The only outlier involves Oid measurements
from~\cite{Razo-Mejia2018}, and as the authors of~\cite{Razo-Mejia2018} note,
this is a difficult measurement of low fluorescence signal against high
background since Oid represses so strongly. We are therefore inclined to regard
the failure of this point to fall near the diagonal as a testament to the
difficulty of the measurement and not as a failure of our theory. \mmnote{Does
this seem a fair characterization?   - RP: thanks, that was very helpful for
me.}

@@

Since we do not have access to repressor copy number for either the
single-cell mRNA data from~\cite{Jones2014} or the single-molecule data
from~\cite{Hammar2014}, we cannot make an apples-to-apples comparison of
association rates $k_R^+$. Further, while Hammar et.\ al.\ directly measure the
dissociation rates $k_R^-$, our inference procedure returns $k_R^-/\gamma$,
i.e., the repressor dissociation rate nondimensionalized by the mRNA degradation
rate $\gamma$. So to make the comparison, we must make an assumption for the
value of $\gamma$ since it was not directly measured. For most mRNAs in
\textit{E.\ coli}, quoted values for the typical mRNA lifetime $\gamma^{-1}$
range between about 2.5~min~\cite{Chen2015} to 8~min. We chose $\gamma^{-1} =
3$~min and $\gamma^{-1} = 5$~min as representative values and plot a comparison
of $k_{O1}^-$ and $k_{Oid}^-$ from our inference with corresponding values
reported in~\cite{Hammar2014} for both these choices of $\gamma$.

The degree of quantitative agreement in the lower panel of
Figure~\ref{fig4:repressed_post_full}B clearly depends on the precise choice of
$\gamma$. Nevertheless we find this comparison very satisfying, when two wildly
different approaches to a measurement of the same quantity yield broadly
compatible results. We emphasize the agreement between our rates and the rates
reported in~\cite{Hammar2014} for any reasonable $\gamma$: values differ by at
most a factor of 2 and possibly agree to within our uncertainties of 10-20\%.
From this we feel confident asserting that the parameters we have inferred from
Jones et.\ al.'s single-cell mRNA counts data do in fact correspond to repressor
binding and unbinding rates, and therefore our conclusions on the agreement of
these rates with binding energies from~\cite{Garcia2011a}
and~\cite{Razo-Mejia2018} are valid.

@@

The details are deferred to Appendix~\ref{sec:bayesian}, and here we only
attempt to summarize the intuitive ideas, as detailed at greater length by
Jaynes~\cite{Jaynes2003} as well as Gelman and
coauthors~\cite{Gelman2013,Gelman2013a}. From our samples of the posterior
distribution, plotted in Figure~\ref{fig4:repressed_post_full}A, we generate
many replicate data using a random number generator. In
Figure~\ref{fig4:repressed_post_full}C, we plot empirical cumulative
distribution functions of the middle 95\% quantiles of these replicate data with
the actual experimental data from Jones et.\ al.~\cite{Jones2014} overlaid,
covering all ten experimental conditions spanning repressor binding sites and
copy numbers (as well as the constitutive baseline UV5).

The purpose of Figure~\ref{fig4:repressed_post_full}C is simply a graphical,
qualitative assessment of the model: do the experimental data systematically
disagree with the simulated data, which would suggest that our model is missing
important features? A further question is not just whether there is a detectable
difference between simulated and experimental data, but whether this difference
is likely to materially affect the conclusions we draw from the posterior in
Figure~\ref{fig4:repressed_post_full}A. More rigorous and quantitative
statistical tests are possible~\cite{Gelman2013}, but their quantitativeness
does not necessarily make them more useful. As stated in~\cite{Gelman2013a}, we
often find this graphical comparison more enlightening because it better engages
our intuition for the model, not merely telling \textit{if} the model is wrong
but suggesting \textit{how} the model may be incomplete.

Our broad brush takeaway from Figure~\ref{fig4:repressed_post_full}C is overall
of good agreement. There some oddities, in particular the long tails in the data
for Oid, 1~ng/mL, and O2, 0.5~ng/mL. The latter is especially odd since it
extends beyond the tail of the unregulated UV5 distribution. This is a
relatively small number of cells, however, so whether this is a peculiarity of
the experimental data, a statistical fluke of small numbers, or a real
biological effect is unclear. It is conceivable that there is some very slow
timescale switching dynamics that could cause this bimodality, although it is
unclear why it would only appear for specific repressor copy numbers. There is
also a small offset between experiment and simulation for O2 at the higher
repressor copy numbers, especially at 2~ng/mL. We suspect this may be caused by
the very large number of zero mRNA counts for Oid, 2~ng/mL skewing it's partner
datasets through the shared association rate, and so we are not too concerned by
this discrepancy either. None of these fairly minor quibbles cause us to
seriously doubt the overall correctness of our model, which further validates
its use to compare the equilibrium models' binding energies to the
nonequilibrium models' repressor kinetics, as we originally set out to do.
